### 2019-11-09 simple_ml/example/report1/paper.tex

\documentclass[journal, a4paper]{IEEEtran}

% some very useful LaTeX packages include:

%\usepackage{cite}      % Written by Donald Arseneau
                        % V1.6 and later of IEEEtran pre-defines the format
                        % of the cite.sty package \cite{} output to follow
                        % that of IEEE. Loading the cite package will
                        % result in citation numbers being automatically
                        % sorted and properly "ranged". i.e.,
                        % [1], [9], [2], [7], [5], [6]
                        % (without using cite.sty)
                        % will become:
                        % [1], [2], [5]--[7], [9] (using cite.sty)
                        % cite.sty's \cite will automatically add leading
                        % space, if needed. Use cite.sty's noadjust option
                        % (cite.sty V3.8 and later) if you want to turn this
                        % off. cite.sty is already installed on most LaTeX
                        % systems. The latest version can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/contrib/supported/cite/

\usepackage{graphicx}   % Written by David Carlisle and Sebastian Rahtz
                        % Required if you want graphics, photos, etc.
                        % graphicx.sty is already installed on most LaTeX
                        % systems. The latest version and documentation can
                        % be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/required/graphics/
                        % Another good source of documentation is "Using
                        % Imported Graphics in LaTeX2e" by Keith Reckdahl
                        % which can be found as esplatex.ps and epslatex.pdf
                        % at: http://www.ctan.org/tex-archive/info/

%\usepackage{psfrag}    % Written by Craig Barratt, Michael C. Grant,
                        % and David Carlisle
                        % This package allows you to substitute LaTeX
                        % commands for text in imported EPS graphic files.
                        % In this way, LaTeX symbols can be placed into
                        % graphics that have been generated by other
                        % applications. You must use latex->dvips->ps2pdf
                        % workflow (not direct pdf output from pdflatex) if
                        % you wish to use this capability because it works
                        % via some PostScript tricks. Alternatively, the
                        % graphics could be processed as separate files via
                        % psfrag and dvips, then converted to PDF for
                        % inclusion in the main file which uses pdflatex.
                        % Docs are in "The PSfrag System" by Michael C. Grant
                        % and David Carlisle. There is also some information
                        % about using psfrag in "Using Imported Graphics in
                        % LaTeX2e" by Keith Reckdahl which documents the
                        % graphicx package (see above). The psfrag package
                        % and documentation can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/contrib/supported/psfrag/

%\usepackage{subfigure} % Written by Steven Douglas Cochran
                        % This package makes it easy to put subfigures
                        % in your figures. i.e., "figure 1a and 1b"
                        % Docs are in "Using Imported Graphics in LaTeX2e"
                        % by Keith Reckdahl which also documents the graphicx
                        % package (see above). subfigure.sty is already
                        % installed on most LaTeX systems. The latest version
                        % and documentation can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/contrib/supported/subfigure/

\usepackage{url}        % Written by Donald Arseneau
                        % Provides better support for handling and breaking
                        % URLs. url.sty is already installed on most LaTeX
                        % systems. The latest version can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/contrib/other/misc/
                        % Read the url.sty source comments for usage information.

%\usepackage{stfloats}  % Written by Sigitas Tolusis
                        % Gives LaTeX2e the ability to do double column
                        % floats at the bottom of the page as well as the top.
                        % (e.g., "\begin{figure*}[!b]" is not normally
                        % possible in LaTeX2e). This is an invasive package
                        % which rewrites many portions of the LaTeX2e output
                        % routines. It may not work with other packages that
                        % modify the LaTeX2e output routine and/or with other
                        % versions of LaTeX. The latest version and
                        % documentation can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/contrib/supported/sttools/
                        % Documentation is contained in the stfloats.sty
                        % comments as well as in the presfull.pdf file.
                        % Do not use the stfloats baselinefloat ability as
                        % IEEE does not allow \baselineskip to stretch.
                        % Authors submitting work to the IEEE should note
                        % that IEEE rarely uses double column equations and
                        % that authors should try to avoid such use.
                        % Do not be tempted to use the cuted.sty or
                        % midfloat.sty package (by the same author) as IEEE
                        % does not format its papers in such ways.

\usepackage{amsmath}    % From the American Mathematical Society
                        % A popular package that provides many helpful commands
                        % for dealing with mathematics. Note that the AMSmath
                        % package sets \interdisplaylinepenalty to 10000 thus
                        % preventing page breaks from occurring within multiline
                        % equations. Use:
%\interdisplaylinepenalty=2500
                        % after loading amsmath to restore such page breaks
                        % as IEEEtran.cls normally does. amsmath.sty is already
                        % installed on most LaTeX systems. The latest version
                        % and documentation can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

\makeatletter
\newif\if@restonecol
\makeatother
\let\algorithm\relax
\let\endalgorithm\relax
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}%[ruled,vlined]{
\usepackage{algpseudocode}
\usepackage{amsmath}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}  % Use Input in the format of Algorithm
\renewcommand{\algorithmicensure}{\textbf{Output:}} % Use Output in the format of Algorithm 
% Other popular packages for formatting tables and equations include:

%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty which improves the
% LaTeX2e array and tabular environments to provide better appearances and
% additional user controls. array.sty is already installed on most systems.
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

% V1.6 of IEEEtran contains the IEEEeqnarray family of commands that can
% be used to generate multiline equations as well as matrices, tables, etc.

% Also of notable interest:
% Scott Pakin's eqparbox package for creating (automatically sized) equal
% width boxes. Available:
% http://www.ctan.org/tex-archive/macros/latex/contrib/supported/eqparbox/

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.

% Your document starts here!
\begin{document}
\begin{titlepage}

\newcommand{\HRule}{\rule{\linewidth}{0.5mm}} % Defines a new command for the horizontal lines, change thickness here

\center % Center everything on the page
 %----------------------------------------------------------------------------------------
%	LOGO SECTION
%----------------------------------------------------------------------------------------

~\\[1cm]
\includegraphics{SCUT.png}\\[2cm] % Include a department/university logo - this will require the graphicx package

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\HRule \\[1cm]
{ \huge \bfseries The Experiment Report of \textit{Machine Learning} }\\[0.6cm] % Title of your document
\HRule \\[2cm]
%----------------------------------------------------------------------------------------
%	HEADING SECTIONS
%----------------------------------------------------------------------------------------

\textsc{\LARGE \textbf{School:} School of Software Engineering}\\[1cm]
\textsc{\LARGE \textbf{Subject:} Software Engineering}\\[2cm]

%----------------------------------------------------------------------------------------
%	AUTHOR SECTION
%----------------------------------------------------------------------------------------

\begin{minipage}{0.4\textwidth}
\begin{flushleft} \large
\emph{Author:}\\
Lizhao Liu % Your name
\end{flushleft}
\end{minipage}
~
\begin{minipage}{0.4\textwidth}
\begin{flushright} \large
\emph{Supervisor:} \\
Mingkui Tan % Supervisor's Name
\end{flushright}
\end{minipage}\\[2cm]
~
\begin{minipage}{0.4\textwidth}
\begin{flushleft} \large
\emph{Student ID:}\\
201730683109
\end{flushleft}
\end{minipage}
~
\begin{minipage}{0.4\textwidth}
\begin{flushright} \large
\emph{Grade:} \\
Undergraduate
\end{flushright}
\end{minipage}\\[2cm]

% If you don't want a supervisor, uncomment the two lines below and remove the section above
%\Large \emph{Author:}\\
%John \textsc{Smith}\\[3cm] % Your name

%----------------------------------------------------------------------------------------
%	DATE SECTION
%----------------------------------------------------------------------------------------

{\large \today}\\[2cm] % Date, change the \today to a set date if you want to be precise

%----------------------------------------------------------------------------------------

\vfill % Fill the rest of the page with whitespace

\end{titlepage}

% Define document title and author
	\title{Linear Regression, Ridge Regression \\and Gradient Descent}
	\maketitle

% Write abstract here
\begin{abstract}
In this report, we solve the house price prediction problem in Housing Dataset by using Linear Regression, Ridge Regression by closed form solution, and even Linear Regression optimized by using Gradient Descent and its variants e.g. Stochastic Gradient Descent and Mini Batch Gradient Descent. \\
We perform experiments on four aspects: \\
1. Comparing the magnitude of $W$ between linear regression and ridge regression optimized by closed form solution. \\
2. Comparision between Gradient Descent and its variants in terms of convergence and time-cost. \\
3. Tuning learning rate of Mini-Batch Gradient Descent. \\
4. Comparing the magnitude of $W$ between linear regression optimized by closed form solution, ridge regression and linear regression optimized by mini-batch Gradient Descent. \\
\end{abstract}

% Each section begins with a \section{title} command
\section{Introduction}
	% \PARstart{}{} creates a tall first letter for this first paragraph
\PARstart{L}{inear} Regression is the core of machine learning. Based on Linear Regression, many machine learning algorithm, such as Ridge Regression, Logistics Regression, SVM are developed. Even in many deep learning model, such as Covolutional Neural Network(CNN), Long Short Term Memory(LSTM), Gated Recurrent Unit(GRU), linear regression(or Linear Layer) is the basic component. So it is very important to fully exploit the insight of Linear Regression. To achieve that goal, In this experiment, we conduct many experiments on linear regression optimized by closed form solution and gradient descent, ridge regression by solving the house price predicting problem.

% Main Part
\section{Methods and Theory}
In this part, we first define the Linear Regression, Mean Squared Error(MES) Loss function. Then we solve the closed form solution of Linear Regression and its variant Ridge Regression. At last we define gradient descent algorithm, and its variants schocastic gradient descent and mini batch gradient descent.
\subsection{Linear Regression}
Given dataset $ D=\{x_{i}, y_{i}\}_{i=1}^{m}$, where $x_{i} \in \mathbf{R}^{n}$ and $y_{i} \in \mathbf{R}$.
Liner Regression Model is parameterized by $(W, b)$, where $W \in \mathbf{R}^{n}$ and $b \in \mathbf{R}$ and the form of it is:
 \begin{center}
 	\begin{equation}
 	y_{i} = x_{i}^{T}W + b
 	\end{equation}
 \end{center} 
We can write it in vertorized form:
 \begin{center}
	\begin{equation}
	y = X^{T}W
	\end{equation}
\end{center} 
where \\
$\mathbf {y} ={\begin{pmatrix}y_{1}\\y_{2}\\\vdots \\y_{m}\end{pmatrix}},\quad 
{\displaystyle X={\begin{pmatrix}1&\mathbf{x} _{1}^{\mathsf {T}}\\1&\mathbf {x} _{2}^{\mathsf {T}}\\\vdots&\vdots \\1&\mathbf {x} _{m}^{\mathsf {T}}\end{pmatrix}}={\begin{pmatrix}1&x_{11}&\cdots &x_{1n}\\1&x_{21}&\cdots &x_{2n}\\\vdots &\vdots &\ddots &\vdots \\1&x_{m1}&\cdots &x_{mn}\end{pmatrix}},} \\
{\displaystyle {\boldsymbol {W }}={\begin{pmatrix}b \\W_{1}\\W_{2}\\\vdots \\W_{n}\end{pmatrix}}.}$
\subsection{Mean Squared Error}
Given the prediction of Linear Model $\hat{y} = X^{T}W $ and the Ground truth $y$, the MSE loss function is:
 \begin{center}
	\begin{equation}
	L(\hat{y}, y) = \frac{1}{2m} \Sigma_{i=1}^{m}(\hat{y}_i, y_i) 
				= \frac{1}{2m}(\hat{y} - y)^{T}(\hat{y} - y)
	\end{equation}
\end{center} 
\subsection{Closed form Solution}
The best parameter $W^{*}$ can be obtained by \\
\begin{equation}
 W^{*} = \arg\min_{W}L(X^{T}W, y)  
\end{equation}
So, we set 
\begin{equation}
\frac{\partial L(\hat{y}, y)}{\partial W} = 0 \\
\end{equation}
We can get 
\begin{equation}
W^{*} = (X^{T}X)^{-1}X^{T}y
\end{equation}
\subsection{Ridge Regression}
In equation 6, when $m > n$, $X^{T}X$ is not invertible. Ridge Regression tackle this problem by adding a small term. So the closed form solution for Ridge Regression is
\begin{equation}
W_{ridge}^{*} = (X^{T}X + \lambda \mathbf{I})^{-1}X^{T}y
\end{equation}
In this way, the loss function for Ridge Regression is
\begin{equation}
L_{ridge}(\hat{y}, y) = \frac{1}{2m}(\hat{y} - y)^{T}(\hat{y} - y)  + \frac{1}{2} W^{T} W
\end{equation}
In another point of view, Ridge Regression can be seen as a regularized form of Linear Regression.
\subsection{Gradient Descent}
Not all problem can be derived a closed form solution, which is often the case in machine learning. So instead of getting the best parameters $W^{*}$ immedeatly, we can update it incremently until $W$ gets closed enough to $W^{*}$. That is the basic idea of Gradient Descent Algorithm. Gradient Descent Algorithm can be found in Algorithm~\ref{alg: Gradient_Descent}. 
\begin{algorithm}

	\label{alg: Gradient_Descent}
	\caption{Gradient Descent}
	\KwIn{Training Dataset $ D=\{x_{i}, y_{i}\}_{i=1}^{m}$, learning rate $\eta $, max iteration $ N$, parameter $W$}
	\KwOut{Optimized parameter $\hat{W}$}
	$t \gets 1$ \\
	\For{$t < N $}
	{
		$\hat{y} \gets X^{T}W $ \\
		$ L(\hat{y}, y) \gets \frac{1}{2m}(\hat{y} - y)^{T}(\hat{y} - y)$ \\
		$\Delta{W} \gets \frac{\partial L}{\partial W}$ \\
		$W \gets W - \eta * \Delta{W}$ \\
	}
	$\hat{W} = W$ \\
	return $\hat{W}$
\end{algorithm}

\subsection{Stochastic Gradient Descent}
Different from gradient descent which updates after it computes the gradient of parameter over all example, stochastic gradient descent updates immediately once it computes parameter's gradient from only one example. Stochastic Gradient Descent Algorithm can be found in Algorithm~\ref{alg: Stochastic_Gradient_Descent}. It can be seen in the algorithm that, there is another for loop inside the for loop of gradient descent algorithm.
\begin{algorithm}
		\label{alg: Stochastic_Gradient_Descent}
	\caption{Stochastic Gradient Descent}
	\KwIn{Training Dataset $ D=\{x_{i}, y_{i}\}_{i=1}^{m}$, learning rate $\eta $, max iteration $ N$, parameter $W$}
	\KwOut{Optimized parameter $\hat{W}$}
	$t \gets 1$ \\
	\For{$t < N $}
	{
		$ i \gets 0 $ \\
		\For{$ i < m $}
		{
			$\hat{y}_{i} \gets X^{T}_{i}W $ \\
			$ L(\hat{y}_{i}, y_{i}) \gets \frac{1}{2}(\hat{y}_{i} - y_{i})^{T}(\hat{y}_{i} - y_{i})$ \\
			$\Delta{W} \gets \frac{\partial L}{\partial W}$ \\
			$W \gets W - \eta * \Delta{W}$ \\
		}
	}
	$\hat{W} = W$ \\
	return $\hat{W}$
\end{algorithm}

\subsection{Mini Batch Gradient Descent}
Due to the fact that stochastic gradient descent updates the parameter based on only one examples, which it is time-comsuming, mini batch gradient descent updates the parameter based on a batch of examples, typically the batch size is bigger than 1 and smaller than the number of whole examples, e.g. 8, 16, 32... Mini Batch Gradient Descent Algorithm can be found in Algorithm~\ref{alg: Mini_Batch_Gradient_Descent}.
\begin{algorithm}
	\label{alg: Mini_Batch_Gradient_Descent}
	\caption{Mini Batch Gradient Descent}
	\KwIn{Training Dataset $ D=\{x_{i}, y_{i}\}_{i=1}^{m}$, learning rate $\eta $, max iteration $ N$, parameter $W$}
	\KwOut{Optimized parameter $\hat{W}$}
	$t \gets 1$ \\
	\For{$t < N $}
	{
		$ i \gets 0 $ \\
		\For{$ X_{batch}, y_{batch} $ in $D$}
		{
			$\hat{y}_{batch} \gets X^{T}_{batch}W $ \\
			$ L(\hat{y}_{batch}, y_{batch}) \gets \frac{1}{2}(\hat{y}_{batch} - y_{batch})^{T}(\hat{y}_{batch} - y_{batch})$ \\
			$\Delta{W} \gets \frac{\partial L}{\partial W}$ \\
			$W \gets W - \eta * \Delta{W}$ \\
		}
	}
	$\hat{W} = W$ \\
	return $\hat{W}$
\end{algorithm}

\section{Experiments}
\subsection{Dataset}
We conduct all the experiments on housing dataset, which has total 506 examples and each example's form is $(x, y)$, where $x \in \mathbf{R}^{13}$ and $y \in \mathbf{R}$. We split the dataset into three part: 272 training examples, 67 validation examples and 167 test examples. $X$ is already scaled between $-1$ and $1$. And we normalized the $X_{train}, X_{validation}, X_{test} $ by using the mean and variance of $X_{train}$.

\subsection{Implementation}
We implement the Linear Regression, Ridge Regression and Gradien Descent Algorithm using python and mainly rely on the numpy package.

\subsection{Linear Regression and Ridge Regression}
In this section, we conduct the experiment of serveral magtitude of $\lambda$, which leads to serveral $W_{ridge}$. We use $L_2{norm}$ to represent the magtitude of a vector. And we evaluate the performance on housing data's test set. Given $v \in \mathbf{R}^{n}$
\begin{equation}
	L_2norm(v) = \sqrt{\Sigma_{i=1}^{n} v_{i}}
\end{equation} \par
From Table~\ref{tab:Ridge_Regression}, we can see that: \par
First, as the magtitude of lambda becomes bigger, both $MSE_{train}$ and $MSE_{val}$ become bigger. \par
Second, as the magtitude of lambda becomes bigger, the magtitude of $W_{ridge}$ become smaller. \par
The first observation is not often the cases. But the second observation can be explained that as the regularizer $\lambda$ become bigger, the $W$ is regularized, so that it is smaller. \par
By comparing Table~\ref{tab:Ridge_Regression} and Table~\ref{tab:Linear_Regression}, we can see that, the performance and the weight magtitude is almost the same as the $\lambda$ is small, but the difference become larger as $\lambda$ grows.
\begin{table}[!hbt]
	% Center the table
	\begin{center}
		% Title of the table
		\caption{Different lambdas' impact in terms of error and $W$ magtitude.}
		\label{tab:Ridge_Regression}
		% Table itself: here we have two columns which are centered and have lines to the left, right and in the middle: |c|c|
		\begin{tabular}{|c|c|c|c|c|c|c|c|}
			% To create a horizontal line, type \hline
			\hline
			% To end a column type &
			% For a linebreak type \\
			$log_{10}^{\lambda}$ & -2 & -1 & 0 & 1 & 2 & 3 \\
			\hline
			$MSE_{train}$   & 10.23 & 10.23 & 10.24 & 10.63 & 30.37 & 181.35   \\
			\hline
			$MSE_{val}$  & 17.08 & 17.09 & 17.16 & 18.20 & 43.92 & 214.55   \\
			\hline
			$L_2norm(W_{ridge})$  & 0.83 & 0.82 & 0.81 & 0.72 & 0.50 & 0.38 \\
			\hline
		\end{tabular}
	\end{center}
\end{table}

\begin{table}[!hbt]
	% Center the table
	\begin{center}
		% Title of the table
		\caption{Linear Regression Error and $W$ magtitude.}
		\label{tab:Linear_Regression}
		% Table itself: here we have two columns which are centered and have lines to the left, right and in the middle: |c|c|
		\begin{tabular}{|c|c|}
			% To create a horizontal line, type \hline
			\hline
			$MSE_{train}$ & 10.23    \\
			\hline
			$MSE_{val}$ & 17.08    \\
			\hline
			$L_2norm(W)$ & 0.83  \\
			\hline
		\end{tabular}
	\end{center}
\end{table}

\subsection{Gradient Descent, Stochastic Gradient Descent and Mini Batch Gradient Descent}
In this section, we conduct experiments on optimizing the linear regression model by using three gradient descent algorrithms, and compare them in terms of convergence and time cost.

\section{Conclusion}
	This section summarizes the paper. In our experiments, you can also write your gains and inspirations in here.

% Your document ends here!

\end{document}